---
title: "Social network analysis of lastfm data using graph"
author: "JAYANTI TRIVEDI"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
---

```{r}
suppressPackageStartupMessages({
  library(purrr)
  library(tidyr)
  library(readr)
  library(igraph)
  library(ggplot2)
  library(dplyr)
})

set.seed(123456)
```

# Social network analysis of lastfm data using graph

## Loads the CSV file with a subset of last.fm social network into an igraph object

## Pre-processes and manipulates data to bring it into analyzable structure

```{r}
# Function for fixing ages (it is real data -- sometimes the age is messed up)
limit_age_ <- function(x) ifelse(x>10 & x<80, x, NA)

# Load edge list dataset
# For col_types I specify character type for both net_user and net_friend
# since they are not used as numbers, they are used as IDs -- character strings
edge_df <- read_csv("lastfm.csv.gz",
                    col_types = "ccdcccdc") %>%
  mutate(age = limit_age_(age),
         friend_age  = limit_age_(friend_age))

# Display
options(dplyr.width = Inf) # Display all columns
head(edge_df)

# Compute vertex attribute data.frame
v_attr_df <- edge_df %>% 
  select(net_user,age,gender,country) %>%
  # combine net_user demographics and net_friend demographics
  # into one long table
  dplyr::union(
    edge_df %>% 
      select(net_friend,friend_age,friend_gender,friend_country) %>%
      rename(net_user = net_friend,
             age = friend_age,
             gender = friend_gender,
             country = friend_country)
  ) %>%
  group_by(net_user) %>%
  # summarise each of the variables by net user
  # duplicates resolved by MAX() for all variables
  summarise_each_(
    funs = funs(max),
    vars = c("age","gender","country")) %>%
  mutate(gender = factor(gender),
         country = factor(country))

# Display vertex attributes
summary(v_attr_df)

# Import into igraph object
g1 <- graph_from_data_frame(
    # all edges
    d = edge_df %>% select(net_user,net_friend),
    # all computed vertex attribute
    vertices = v_attr_df
  ) %>%
  # collapse convert directed to non-directed
  as.undirected(mode = "collapse")
```

## Top 5 countries of friends

```{r}
TopN <- 5L

friends_country <- V(g1)[unique(edge_df$net_friend)]$country

q2 <- tail(names(sort(table(friends_country))), TopN)

q2
```


## Plots friendships between top 5 countries rather than between users
### Thickness of edges represents friendships
### Edge labels represent total number of friendships, like if country A has 54,000 friends with country B then edge labels display 54

```{r}
friend_df = filter(v_attr_df, net_user %in% edge_df$net_friend)

final_friend_df = inner_join(edge_df, friend_df, 
                             by = c("net_friend" = "net_user"))

ego_df = filter(v_attr_df, net_user %in% edge_df$net_user)

final_df <- inner_join(final_friend_df, ego_df, by = "net_user")

all5 <- final_df %>% 
  filter(country %in% q2 & country.y %in% q2) %>% 
  select(country,country.y)

all5$friendship = 1

all5_df <- all5 %>% 
  group_by(country,country.y) %>% 
  summarise(friendship=sum(friendship))

gg <- graph_from_data_frame(all5_df, directed = TRUE)
E(gg)$weight <- E(gg)$friendship

g3 <- as.undirected(gg, mode = "collapse")

V(g3)$name <- c("BR","PO","RU","UK","US")
V(g3)$size <- 30L
E(g3)$curved <- TRUE
plot(g3, edge.label = round(E(g3)$weight/1000,1), 
     edge.width = E(g3)$weight/5600,
     edge.curved=0.2)
```
